Reasoning about a Capability Machine with Local Capabilities

### 2017-06-23 popl-2018/paper.tex

Our logical relation is a unary, step-indexed Kripke logical relation
with recursive
worlds \citep{pitts_operational_1998,Appel:2001:IMR:504709.504712,Ahmed2004semantics,Birkedal:2011:SKM:1926385.1926401},
closely related to the one used by \citet{Devriese:2016ObjCap} to formulate
capability safety in a high-level JavaScript-like lambda calculus. Our
Fundamental Theorem is similar to theirs and expresses
capability safety of the capability machine. Because we are not interested in
externally observable side-effects (like console output or memory access
traces), we do not require their notion of effect parametricity. Our logical
relation uses several ideas from previous work, like Kripke worlds with regions
containing state transition systems \citep{Ahmed:popl09}, public/private future
worlds \citep{Dreyer:jfp12} (see Section~\ref{sec:discussion} for a discussion), and
biorthogonality \citep{pitts_operational_1998,benton_biorthogonality_2009-1,Hur:2011:KLR:1926385.1926402}.
Swasey et. al.\ have recently developed a \emph{logic}, OCPL,  for verification of
object capability patterns \citep{swasey:2017}. The logic is based on
Iris \citep{iris,iris2,iris3}, a state of the art higher-order
concurrent separation logic and is formalized in Coq, building on the
Iris Proof Mode for Coq \citep{ipm}.

>>>>

Our logical relation is a unary, step-indexed Kripke logical relation with
recursive worlds
\citep{pitts_operational_1998,Appel:2001:IMR:504709.504712,Ahmed2004semantics,Birkedal:2011:SKM:1926385.1926401},
closely related to the one used by \citet{Devriese:2016ObjCap} to formulate
capability safety in a high-level JavaScript-like lambda calculus. Our
Fundamental Theorem is similar to theirs and expresses capability safety of the
capability machine. Because we are not interested in externally observable
side-effects (like console output or memory access traces), we do not require
their notion of effect parametricity. Our logical relation uses several ideas
from previous work, like Kripke worlds with regions containing state transition
systems \citep{Ahmed:popl09}, public/private future worlds \citep{Dreyer:jfp12}
(see Section~\ref{sec:discussion} for a discussion), and biorthogonality
\citep{pitts_operational_1998,benton_biorthogonality_2009-1,Hur:2011:KLR:1926385.1926402}.

Swasey et. al.\ have recently developed a \emph{logic}, OCPL, for verification
of object capability patterns~\citep{swasey:2017}. The logic is based on
Iris~\citep{iris,iris2,iris3}, a state of the art higher-order concurrent
separation logic and is formalized in Coq, building on the Iris Proof Mode for
Coq~\citep{ipm}. OCPL gives a more abstract and modular way of proving
capability safety for a lambda-calculus (with concurrency) compared to the
earlier work by \citet{Devriese:2016ObjCap}. In the future we would also like to
investigate a new program logic for reasoning about capability safety for our
capability machine model. We think Iris would also be a natural starting point
for such an endeavour, since Iris is really a framework, which can be
instantiated to different programming languages. OCPL was able to leverage
existing Iris specifications for the high-level language; for our capability
machine model, however, it would be necessary to devise new kinds of
specifications for our low-level programs with unstructured control-flow. It is
likely that we could get inspiration from earlier work on logics for assembly
programming languages, such as XCAP~\citep{xcap}.
